---
title: "Final Project"
author: "Ryan Zomorrodi"
date:  last-modified
categories: ["Final Project"]
format: 
  html:
    toc: true
    code-fold: true
    embed-resources: true
execute:
  cache: true
  warning: false
  error: false
csl: ieee.csl

references:
- id: huynh_estimated_2024
  type: article-journal
  author:
  - family: Huynh
    given: Benjamin Q.
  - family: Chin
    given: Elizabeth T.
  - family: Kiang
    given: Mathew V.
  issued:
  - year: 2024
    month: 3
  title: Estimated Childhood Lead Exposure From Drinking Water in Chicago
  container-title: JAMA pediatrics
  page: e240133
  abstract: 'IMPORTANCE: There is no level of lead in drinking water considered to
    be safe, yet lead service lines are still commonly used in water systems across
    the US. OBJECTIVE: To identify the extent of lead-contaminated drinking water
    in Chicago, Illinois, and model its impact on children younger than 6 years. DESIGN,
    SETTING, AND PARTICIPANTS: For this cross-sectional study, a retrospective assessment
    was performed of lead exposure based on household tests collected from January
    2016 to September 2023. Tests were obtained from households in Chicago that registered
    for a free self-administered testing service for lead exposure. Machine learning
    and microsimulation were used to estimate citywide childhood lead exposure. EXPOSURE:
    Lead-contaminated drinking water, measured in parts per billion. MAIN OUTCOMES
    AND MEASURES: Number of children younger than 6 years exposed to lead-contaminated
    water. RESULTS: A total of 38 385 household lead tests were collected. An estimated
    68% (95% uncertainty interval, 66%-69%) of children younger than 6 years were
    exposed to lead-contaminated water, corresponding to 129 000 children (95% uncertainty
    interval, 128 000-131 000 children). Ten-percentage-point increases in block-level
    Black and Hispanic populations were associated with 3% (95% CI, 2%-3%) and 6%
    (95% CI, 5%-7%) decreases in odds of being tested for lead and 4% (95% CI, 3%-6%)
    and 11% (95% CI, 10%-13%) increases in having lead-contaminated drinking water,
    respectively. CONCLUSIONS AND RELEVANCE: These findings indicate that childhood
    lead exposure is widespread in Chicago, and racial inequities are present in both
    testing rates and exposure levels. Machine learning may assist in preliminary
    screening for lead exposure, and efforts to remediate the effects of environmental
    racism should involve improving outreach for and access to lead testing services.'
  DOI: 10.1001/jamapediatrics.2024.0133
  ISSN: 2168-6211
  PMCID: PMC10949143
  PMID: '38497944'
---

## Introduction

The following project is a replication attempt of the machine learning model described in Huynh et al. @huynh_estimated_2024 predicting lead exposure from drinking water within the city of Chicago. Lead-contaminated drinking water at the block level was defined as a binary variable indicating whether the majority of tests within a block have at least 1 ppb lead concentration. According to their model, approximately 75% of blocks are estimated to have lead-contaminated drinking water. The greatest predictors of lead-contaminated drinking water were geographic areas, population at the block, and number of buildings.

Huynh et al. (and by extension, I) used the following data sources:

<details class="code-fold">
<summary>Data Sources</summary>
+---------------------------------------------------------------+-----------------------------------------------------------+-------------------------+
| Source                                                        | Measure                                                   | Extent                  |
+===============================================================+===========================================================+=========================+
| City of Chicago Department of Water Management Lead Test Data | Consecutive lead tests (ppb)                              | Anonymized to the block |
+---------------------------------------------------------------+-----------------------------------------------------------+-------------------------+
| Census                                                        | Block FIPS                                                | Block                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Population (#)                                            | Block                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Race/ethnicity (#)                                        | Block                   |
|                                                               |                                                           |                         |
|                                                               | -   AIAN                                                  |                         |
|                                                               |                                                           |                         |
|                                                               | -   Asian                                                 |                         |
|                                                               |                                                           |                         |
|                                                               | -   Black                                                 |                         |
|                                                               |                                                           |                         |
|                                                               | -   Hispanic                                              |                         |
|                                                               |                                                           |                         |
|                                                               | -   White                                                 |                         |
+---------------------------------------------------------------+-----------------------------------------------------------+-------------------------+
| American Community Survey                                     | Block FIPS                                                | Block                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Block group FIPS                                          | Block group             |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Population (#)                                            | Block group, tract      |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Race/ethnicity (#)                                        | Block group, tract      |
|                                                               |                                                           |                         |
|                                                               | -   AIAN                                                  |                         |
|                                                               |                                                           |                         |
|                                                               | -   Asian                                                 |                         |
|                                                               |                                                           |                         |
|                                                               | -   Black                                                 |                         |
|                                                               |                                                           |                         |
|                                                               | -   Hispanic                                              |                         |
|                                                               |                                                           |                         |
|                                                               | -   White                                                 |                         |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Housing units (#)                                         | Block group             |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Median house value (\$)                                   | Block group             |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Upper house value (\$)                                    | Block group             |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Lower house value (\$)                                    | Block group             |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Median homeowner costs (\$)                               | Block group             |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Education (#)                                             | Block group             |
|                                                               |                                                           |                         |
|                                                               | -   High school                                           |                         |
|                                                               |                                                           |                         |
|                                                               | -   GED                                                   |                         |
|                                                               |                                                           |                         |
|                                                               | -   \<1 year of college                                   |                         |
|                                                               |                                                           |                         |
|                                                               | -   \> 1 year of college                                  |                         |
|                                                               |                                                           |                         |
|                                                               | -   Associate's degree                                    |                         |
|                                                               |                                                           |                         |
|                                                               | -   Bachelor's degree                                     |                         |
|                                                               |                                                           |                         |
|                                                               | -   Master's degree                                       |                         |
|                                                               |                                                           |                         |
|                                                               | -   Professional School                                   |                         |
|                                                               |                                                           |                         |
|                                                               | -   Doctoral Degree                                       |                         |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Poverty (#)                                               | Block group             |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | English-only speakers (#)                                 | Block group             |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Computer access (#)                                       | Block group             |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Internet access (#)                                       | Block group             |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Complete plumbing facilities (#)                          | Block group             |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Vacant Housing (#)                                        | Block group             |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Owner-occupied (#)                                        | Block group             |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Children under 5 (#)                                      | Block group             |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Children under 10 (#)                                     | Block group             |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Children under 18 (#)                                     | Block group             |
+---------------------------------------------------------------+-----------------------------------------------------------+-------------------------+
| Chicago Building Footprints                                   | Median building age (years)                               | Block                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Building (#)                                              | Block                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Max age (years)                                           | Block                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Mean age (years)                                          | Block                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Built after 1986 (#)                                      | Block                   |
+---------------------------------------------------------------+-----------------------------------------------------------+-------------------------+
| Chicago Health Atlas                                          | Community area                                            | Community area          |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Lead poisoning rate (%)                                   | Community area          |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Historical lead poisoning rate (%)                        | Community area          |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Economic diversity index                                  | Tract                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Hardship index                                            | Tract                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Social vulnerability index                                | Tract                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Major crime (#)                                           | Tract                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Eviction rate (%)                                         | Tract                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Fine particulate matter concentration ($\mu\text{g/m}^3$) | Tract                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Access to food (%)                                        | Tract                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Uninsured rate (%)                                        | Tract                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Per capita income (\$)                                    | Tract                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Cognitive difficulty (%)                                  | Tract                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Disability (%)                                            | Tract                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Crowded housing (%)                                       | Tract                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Rent burdened (%)                                         | Tract                   |
+                                                               +-----------------------------------------------------------+-------------------------+
|                                                               | Vacant housing (%)                                        | Tract                   |
+---------------------------------------------------------------+-----------------------------------------------------------+-------------------------+
</details>



## Data

Libraries
```{r}
library(bonsai)
library(corrr)
library(finetune)
library(future)
library(httr2)
library(jsonlite)
library(leaflet)
library(magrittr)
library(mice)
library(readxl)
library(sf)
library(themis)
library(tidycensus)
library(tidymodels)
library(tidyverse)
library(tigris)
```

### Chicago Boundaries

#### Download
Download boundaries at various levels: the city, community areas, and census blocks.
```{r}
#| output: false

chicagoBoundaries <- st_read("https://data.cityofchicago.org/api/geospatial/qqq8-j68g?fourfour=qqq8-j68g&cacheBust=1712775952&date=20240411&accessType=DOWNLOAD&method=export&format=GeoJSON") %>%
    st_transform("EPSG:4269")

chicagoCommunityAreas <- st_read("https://data.cityofchicago.org/api/geospatial/cauq-8yn6?method=export&format=GeoJSON") %>%
    st_transform("EPSG:4269")

cookBlocks <- blocks(state = "IL", county = "Cook") 
```

#### Process
Perform an intersection on census blocks to subset to those blocks that are within Chicago Boundaries. Mutate GEOIDs to create complete block, block group, and tract GEOIDS.
```{r}
#| eval: false

chicagoBlocks <- cookBlocks %>%
    filter(st_intersects(., chicagoBoundaries, sparse = FALSE) %>% unlist()) %>%
    filter(!str_detect(TRACTCE20, "^9900")) %>%
    mutate(GEOID_blk = GEOID20) %>%
    mutate(GEOID_blkGrp = str_sub(GEOID_blk, 1, -4)) %>%
    mutate(GEOID_tract = str_sub(GEOID_blk, 1, -5)) %>% 
    select(starts_with("GEOID_"))
```

### American Community Survey

#### Download
Get metadata of all the relevant American Community Survey data.
```{r}
#| eval: false

census_metadata <- 
    bind_rows(
        load_variables(2022, "acs5"), 
        load_variables(2020, "pl"))

blk_vars <- c(
    race_blk_total = "P2_001N",
    race_blk_aianNH = "P2_007N",
    race_blk_asianNH = "P2_008N",
    race_blk_blackNH = "P2_006N",
    race_blk_hispanic = "P2_002N",
    race_blk_whiteNH = "P2_005N"
)

blkGrp_vars <- c(
    race_blkGrp_total = "B03002_001",
    race_blkGrp_aianNH = "B03002_005",
    race_blkGrp_asianNH = "B03002_006",
    race_blkGrp_blackNH = "B03002_004",
    race_blkGrp_hispanic = "B03002_012",
    race_blkGrp_whiteNH = "B03002_003",
    housingUnits = "B25001_001",
    lowerValue = "B25076_001",
    medValue = "B25077_001",
    upperValue = "B25078_001",
    homeownerCost = "B25088_001",
    education_total = "B15003_001",
    education_highSchool = "B15003_017",
    education_ged = "B15003_018",
    education_lt1yCollege = "B15003_019",
    education_mt1yCollege = "B15003_020",
    education_associate = "B15003_021",
    education_bachelor = "B15003_022",
    education_master = "B15003_023",
    education_professional = "B15003_024",
    education_doctorate = "B15003_025",
    poverty_total = "B17010_001",
    poverty_belowPovertyLvl = "B17010_002", 
    language_total = "B99162_001",
    language_onlyEnglish = "B99162_002",
    computer_total = "B28003_001",
    computer_hasAComp = "B28003_002",
    internet_total = "B28011_001",
    internet_noInternet = "B28011_008",
    plumbing_total = "B25047_001",
    plumbing_complete = "B25047_002",
    vacancy_total = "B25002_001",
    vacancy_vacant = "B25002_003",
    occupied_total = "B25003_001",
    occupied_owner = "B25003_002",
    age_total = "B01001_001",
    age_maleU5 = "B01001_003",
    age_male5to9 = "B01001_004",
    age_male10to14 = "B01001_005",
    age_male15to17 = "B01001_006",
    age_femaleU5 = "B01001_027",
    age_female5to9 = "B01001_028",
    age_female10to14 = "B01001_029",
    age_female15to17 = "B01001_030")

tract_vars <- c(
    race_tract_total = "B03002_001",
    race_tract_aianNH = "B03002_005",
    race_tract_asianNH = "B03002_006",
    race_tract_blackNH = "B03002_004",
    race_tract_hispanic = "B03002_012",
    race_tract_whiteNH = "B03002_003",
    native_total = "B05002_001",
    native_nativeBrn = "B05002_002")

select_vars_metadata <- census_metadata %>%
    filter(name %in% c(blk_vars, blkGrp_vars, tract_vars))
```

Download American Community Survey data for all the relevant variables
```{r}
#| eval: false

blk_data <- get_decennial(geography = "block", variables = blk_vars, state = "IL", county = "Cook", output = "wide")
blkGrp_data <- get_acs(geography = "block group", variables = blkGrp_vars, state = "IL", county = "Cook", output = "wide") 
tract_data <- get_acs(geography = "tract", variables = tract_vars, state = "IL", county = "Cook", output = "wide") 
```

#### Process
Join datasets and utilize multiple imputation by chained equations with random forest as the regression model.
```{r}
#| eval: false

chicagoACS <- blk_data %>%
    select(-NAME) %>%
    rename(GEOID_blk = GEOID) %>%
    mutate(GEOID_blkGrp = str_sub(GEOID_blk, 1, -4)) %>%
    mutate(GEOID_tract = str_sub(GEOID_blk, 1, -5)) %>%
    relocate(GEOID_blkGrp, GEOID_tract, .after = GEOID_blk) %>%
    left_join(
        blkGrp_data %>%
            select(-ends_with("M"), -NAME) %>%
            mice(meth = "rf", seed = 123) %>%
            complete() %>%
            as_tibble(),
        join_by(GEOID_blkGrp == GEOID)) %>%
    left_join(
        tract_data %>%
            select(-NAME), 
        join_by(GEOID_tract == GEOID)) %>%
    mutate(
        age_U5E = age_maleU5E + age_femaleU5E,
        age_5to9E = age_male5to9E + age_female5to9E,
        age_10to17E = age_male10to14E + age_female10to14E + age_male15to17E + age_male15to17E) %>%
    select(-contains("male"), -ends_with("M")) %>%
    right_join(
        chicagoBlocks %>%
            st_drop_geometry()
    )
```

### Chicago Building Footprints

#### Download
Download Chicago Building Footprints dataset to identify the age and number of buildings.
```{r}
#| eval: false

chicagoBuildings <- st_read("https://data.cityofchicago.org/api/geospatial/syp8-uezg?fourfour=syp8-uezg&cacheBust=1712775954&date=20240414&accessType=DOWNLOAD&method=export&format=GeoJSON") %>%
    st_transform("EPSG:4269")
```

#### Process
Remove buildings with empty geometries or empty street names. Spatial join buildings to the block. Summarize building characteristics at the block level. 
```{r}
#| eval: false

sf_use_s2(FALSE)

chicagoBuildingsImp <- chicagoBuildings %>%
    filter(!st_is_empty(.)) %>%
    filter(st_name1 != "") %>%
    mutate(year_built =
        case_when(
            year_built == 0 ~ NA,
            .default = as.integer(year_built)
        )      
    ) %>%
    select(year_built) %>%
    st_join(chicagoBlocks) %>%
    st_drop_geometry() %>%
    left_join(chicagoACS) %>%
    tibble() %>%
    mice(meth = "rf", seed = 234) %>%
    complete() %>%
    as_tibble()

chicagoBldBlk <- chicagoBuildingsImp %>%
    st_drop_geometry() %>%
    group_by(GEOID_blk) %>%
    summarise(
        bldAge_median = 2024 - median(year_built),
        bld_number = n(),
        bldAge_max = 2024 - min(year_built),
        bldAge_mean = 2024 - mean(year_built),
        bldAge_nAft86 = sum(year_built > 1986)
    )
```

### Chicago DWM Lead Tests

#### Download
Download Department of Water Management lead testing data 
```{r}
#| eval: false

pb_test_path <- tempfile()
download.file("https://www.chicagowaterquality.org/DataFiles/wqContent/Results.xlsx", pb_test_path, mode="wb")

chicagoPbTest <- read_excel(path = pb_test_path, skip = 2, sheet = 1) 
```


#### Process
```{r}
#| eval: false

chicagoBuildingAddresses <- chicagoBuildings %>%
    filter(!st_is_empty(.)) %>%
    filter(st_name1 != "") %>%
    st_join(chicagoBlocks) %>%
    st_drop_geometry() %>%
    as_tibble() %>%
    arrange(t_add1) %>%
    mutate(ID = row_number()) %>%
    mutate(
        Address = str_pad(t_add1, 2, pad = "0"),
        Address = str_c(str_sub(Address, end = -3), "XX"),
        Address = str_c(Address, pre_dir1, st_name1, st_type1, sep = " "))

chicagoPbBlk <- chicagoPbTest %>%
    filter(!is.na(`Sample Date`)) %>%
    filter(`1st Draw` != "See Follow-Up Sequential Sampling Table for Results") %>%
    mutate(across(c(`1st Draw`, `2/3 Min`, `5 Min`), ~
        case_when(
            str_detect(.x, "<") ~ NA,
            .default = .x
        )
    )) %>%
    mutate(across(c(`1st Draw`, `2/3 Min`, `5 Min`), ~
        case_when(
            is.na(.x) ~ TRUE,
            .default = FALSE
        ),
        .names = "{.col}_lt1"
    )) %>%
    mutate(ID = 
        map_int(Address, ~
            chicagoBuildingAddresses %>%
                filter(Address == .x) %>%
                slice( (n() + (n() %% 2))/2) %>%
                pluck("ID") %>%
                {ifelse(length(.) == 0, NA, .)}
        )
    ) %>% 
    left_join(chicagoBuildingAddresses) %>%
    group_by(GEOID_blk) %>%
    summarize(PB_gt1Pct = 1 - mean(`1st Draw_lt1`), PB_nTests = n()) %>%
    mutate(PB_majoritygt1 = PB_gt1Pct >= 0.5)
```

### Chicago Health Atlas

```{r}
#| eval: false

chicagoHAtract <- request("https://chicagohealthatlas.org/api/action/download/") %>%
    req_method("POST") %>%
    req_body_raw(r"[{"layer":"tract-2020","topics":"EDX-~2018-2022,HDX-~2018-2022,SVI-~2020,CZM-~2018-2022,EVR-~2018,PMC-~2023,LFA-~2019,UNS-~2018-2022,PCI-~2018-2022,DIV-~2018-2022,DIS-~2018-2022,HTJ-~2018-2022,RBU-~2018-2022,VAC-~2018-2022","state":"","within":"","regions":"","place_filters":"","errors":"se","population":false,"lat_long":false,"counties":false,"documentation":false,"format":"csv","insight":"","benchmark":false}]", "application/x-www-form-urlencoded") %>%
    req_perform() %>%
    resp_body_string() %>%
    {read_csv(., col_names = names(read_csv(., n_max = 0)), cols(GEOID = "c"), skip = 2)} %>%
    select(-Layer, -Name) %>%
    rename(GEOID_tract = GEOID) %>%
    drop_na(!ends_with("se"))


chicagoHAPbCA <- request("https://chicagohealthatlas.org/api/action/download/") %>%
    req_method("POST") %>%
    req_body_raw(r"[{"layer":"neighborhood","topics":"LDPP-~2023,LDPPH-~2016","state":"","within":"","regions":"","place_filters":"","errors":"se","population":false,"lat_long":false,"counties":false,"documentation":false,"format":"csv","insight":"","benchmark":false}]", "application/x-www-form-urlencoded") %>%
    req_perform() %>%
    resp_body_string() %>%
    {read_csv(., col_names = names(read_csv(., n_max = 0)), cols(GEOID = "c"), skip = 2)}

chicagoHAblk <- chicagoBlocks %>%
    st_join(
        left_join(chicagoHAPbCA, chicagoCommunityAreas, join_by(GEOID == area_numbe)) %>%
            st_as_sf(), 
        join = st_covered_by, 
        largest = TRUE) %>%
    rename(GEOID_CA = GEOID) %>%
    left_join(chicagoHAtract) %>%
    st_drop_geometry() %>%
    as_tibble() %>%
    select(-Layer, -Name, -community, -area, -shape_area, -perimeter, -area_num_1, -comarea_id, -comarea, -shape_len) %>%
    drop_na(!ends_with("se"))
```

### Chicago Service Line Inventory

#### Download
Not currently implemented due to time constraints. This an additional data source that I am contemplating using.
```{r}
#| eval: false

resp <- building_blocks %>%
    st_drop_geometry() %>%
    as_tibble() %>%
    filter(year_built != 0) %>%
    slice(1:100) %>% 
    mutate(st_add = str_c(label_hous, pre_dir1, st_name1, st_type1, sep = " ")) %>%
    pull(st_add) %>% 
    map(~
        request("https://sli.chicagowaterquality.org/servicematerialsearch/api/") %>%
            req_method("POST") %>%
            req_body_raw(str_c(r"[{"type":"text","accNum":"-","addr":"]", .x, r"["}]"))
    ) %>%
    req_perform_parallel(on_error = "continue")

resp %>%
    resps_successes() %>%
    keep(map_lgl(., resp_has_body)) %>%
    resps_data(resp_body_json) %>%
    map(~ fromJSON(.x) %>% `[[`(1)) %>%
    map_df(bind_cols)
```

### Joining and Conducting Final Processing
In the end, the following criteria was used to include blocks within the study:  

-   Population greater than 0
-   Number of buildings greater than 0
-   Within tract included in the Chicago Health Atlas

This leaves us with:  

-   33,004 total blocks
-   12,031 blocks with at least one prior DWM test
    -   0.799 blocks have a majority of first draw tests with detected lead levels greater than 1 ppb 

```{r}
#| eval: false

pct_calculator <- function(data, category) {
    mutate(data, across(
        .cols = matches(str_c("^", category, "(.*)(?<!(total))E$"), perl = TRUE), 
        .fn = ~ .x / !!sym(str_c(category, "_totalE")),
        .names = "{.col}Pct")) %>%
    select(-matches(str_c("^", category, "(.*)(?<!(total))E$"), perl = TRUE))
}

data <- chicagoACS %>%
    left_join(chicagoPbBlk) %>%
    inner_join(chicagoBldBlk) %>%
    inner_join(chicagoHAblk) %>%
    select(-ends_with("se")) %>%
    relocate(GEOID_CA, PB_gt1Pct, PB_nTests, PB_majoritygt1, .after = GEOID_tract) %>%
    rename_with(.cols = starts_with("race_blk_"), .fn = ~ str_c(.x, "E")) %>%
    filter(race_blk_totalE > 0) %>%
    pct_calculator("race_blk") %>%
    pct_calculator("race_blkGrp") %>%
    pct_calculator("race_tract") %>%
    pct_calculator("education") %>%
    pct_calculator("poverty") %>%
    pct_calculator("computer") %>%
    pct_calculator("internet") %>%
    pct_calculator("plumbing") %>%
    pct_calculator("vacancy") %>%
    pct_calculator("occupied") %>%
    pct_calculator("age") %>%
    pct_calculator("language") %>%
    pct_calculator("native") %>%
    drop_na(-starts_with("PB")) %>%
    select(-matches("^(?!race)(.*)_totalE$", perl = TRUE)) %>%
    rename_with(.cols = starts_with("race"), ~ 
        str_replace(.x, "race", "population") %>%
            str_replace("_totalE", "E")
    )

data %>%
    summarize(across(everything(), ~ sum(is.na(.x)))) %>%
    View()
```

```{r}
#| include: false
#| eval: false

save(data, file = "../data/lead_data.RData")
```

```{r}
#| include: false

load("../data/lead_data.RData")
```

## Exploratory Data Analysis

Tracts that test more tend to have a higher percent of water tests with greater than 1 PPB lead even when adjusting for population size. 

::: panel-tabset
## Lead Positive Percentage by Number of Tests
```{r}
data %>%
    group_by(GEOID_tract) %>%
    summarize(
        n = sum(PB_nTests, na.rm = TRUE),
        gt1pct = sum(PB_nTests * PB_gt1Pct, na.rm = TRUE)/ n
    ) %>% 
    group_by(n, gt1pct) %>%
    summarize(nTract = n()) %>% 
    ggplot(aes(x = n, y = gt1pct, color = nTract, size = nTract)) +
        geom_point() +
        labs(
            title = "Number of Tests vs Percent of Tests with Greater than 1 \nPPB Lead Detected by Tract",
            x = "Tests (#)",
            y = "Tests with Greater than 1 PPB Lead Detected (%)") +
        scale_color_gradient(low = "steelblue1", high = "black") +
        scale_y_continuous(labels = scales::percent) +
        theme(
            plot.title = element_text(hjust = 0.5),
            legend.position="none"
        )
```

## Lead Positive Percentage by Number of Tests Per Person
```{r}
data %>%
    group_by(GEOID_tract) %>%
    summarize(
        nPer100People = 100 *sum(PB_nTests, na.rm = TRUE) / population_tractE[1],
        gt1pct = sum(PB_nTests * PB_gt1Pct, na.rm = TRUE)/  sum(PB_nTests, na.rm = TRUE)
    ) %>% 
    ggplot(aes(x = nPer100People, y = gt1pct)) +
        geom_point(color = "steelblue1") +
        labs(
            title = "Number of Tests per 100 Residents vs Percent of Tests with Greater than 1 \nPPB Lead Detected by Tract",
            x = "Tests Per 100 Residents (#)",
            y = "Tests with Greater than 1 PPB Lead Detected (%)") +
        scale_y_continuous(labels = scales::percent) +
        theme(
            plot.title = element_text(hjust = 0.5),
            legend.position="none"
        )
```
:::

We can see that residents on the north side have disproportionately low positive lead rates when compared to that of the rest of Chicago

::: panel-tabset
## Positive Test Percentage by Block Group
```{r}
#| fig-height: 10

data %>%
    group_by(GEOID_blkGrp) %>%
    summarize(
        n = sum(PB_nTests, na.rm = TRUE),
        gt1pct = sum(PB_nTests * PB_gt1Pct, na.rm = TRUE) / n
    ) %>% 
    left_join(block_groups(state = "IL", county = "Cook", progress_bar = FALSE), join_by(GEOID_blkGrp == GEOID)) %>%
    st_as_sf() %$%
        {leaflet(.) %>%
        addProviderTiles(providers$CartoDB.Positron) %>% 
        addPolygons(
            fillColor = ~ colorNumeric("Blues", gt1pct)(gt1pct),
            fillOpacity = 0.7,
            weight = 0,
            highlightOptions = highlightOptions(
                color = "Black",
                fillOpacity = 1,
                bringToFront = T),
            label = str_glue_data(., "<strong>{GEOID_blkGrp}</strong><br>Percent of Tests: {round(gt1pct,2)}<br/>") %>% 
                    map(htmltools::HTML),
            labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "12px",
                direction = "auto")
        ) %>%
        addLegend(
            pal = colorNumeric("Blues", gt1pct),
            values = gt1pct,
            opacity = 0.7,
            title = "Tests with Greater than 1 PPB Lead <br> Detected (%) By Block Group",
            position = "topright",
            na.label = "Insufficient Data",
            labFormat = labelFormat(
                prefix = " ", suffix = " %",
                transform = function(x) 100 * x)
        )} %>%
        htmlwidgets::prependContent(htmltools::tags$style(type = "text/css", "div.info.legend.leaflet-control br {clear: both;}"))
```

## Positive Test Percentage by Tract
```{r}
#| fig-height: 10

data %>%
    group_by(GEOID_tract) %>%
    summarize(
        n = sum(PB_nTests, na.rm = TRUE),
        gt1pct = sum(PB_nTests * PB_gt1Pct, na.rm = TRUE) / n
    ) %>% 
    left_join(tracts(state = "IL", county = "Cook", progress_bar = FALSE), join_by(GEOID_tract == GEOID)) %>%
    st_as_sf() %$%
        {leaflet(.) %>%
        addProviderTiles(providers$CartoDB.Positron) %>% 
        addPolygons(
            fillColor = ~ colorNumeric("Blues", gt1pct)(gt1pct),
            fillOpacity = 0.7,
            weight = 0,
            highlightOptions = highlightOptions(
                color = "Black",
                fillOpacity = 1,
                bringToFront = T),
            label = str_glue_data(., "<strong>{GEOID_tract}</strong><br>Percent of Tests: {round(gt1pct,2)}<br/>") %>% 
                    map(htmltools::HTML),
            labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "12px",
                direction = "auto")
        ) %>%
        addLegend(
            pal = colorNumeric("Blues", gt1pct),
            values = gt1pct,
            opacity = 0.7,
            title = "Tests with Greater than 1 PPB Lead <br> Detected (%) By Tract",
            position = "topright",
            na.label = "Insufficient Data",
            labFormat = labelFormat(
                prefix = " ", suffix = " %",
                transform = function(x) 100 * x)
        )} %>%
        htmlwidgets::prependContent(htmltools::tags$style(type = "text/css", "div.info.legend.leaflet-control br {clear: both;}"))
```

## Positive Test Percentage by Community Area
```{r}
#| fig-height: 10

data %>%
    group_by(GEOID_CA) %>%
    summarize(
        n = sum(PB_nTests, na.rm = TRUE),
        gt1pct = sum(PB_nTests * PB_gt1Pct, na.rm = TRUE) / n
    ) %>% 
    left_join(chicagoCommunityAreas, join_by(GEOID_CA == area_numbe)) %>%
    st_as_sf() %$%
        {leaflet(.) %>%
        addProviderTiles(providers$CartoDB.Positron) %>% 
        addPolygons(
            fillColor = ~ colorNumeric("Blues", gt1pct)(gt1pct),
            fillOpacity = 0.7,
            weight = 0,
            highlightOptions = highlightOptions(
                color = "Black",
                fillOpacity = 1,
                bringToFront = T),
            label = str_glue_data(., "<strong>{community}</strong><br>Percent of Tests: {round(gt1pct,2)}<br/>") %>% 
                    map(htmltools::HTML),
            labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "12px",
                direction = "auto")
        ) %>%
        addLegend(
            pal = colorNumeric("Blues", gt1pct),
            values = gt1pct,
            opacity = 0.7,
            title = "Tests with Greater than 1 PPB Lead <br> Detected (%) By Community Area",
            position = "topright",
            na.label = "Insufficient Data",
            labFormat = labelFormat(
                prefix = " ", suffix = " %",
                transform = function(x) 100 * x)
        )} %>%
        htmlwidgets::prependContent(htmltools::tags$style(type = "text/css", "div.info.legend.leaflet-control br {clear: both;}"))
```

:::

::: panel-tabset
## Tests Per Person by Block Group
```{r}
#| fig-height: 10

data %>%
    group_by(GEOID_blkGrp) %>%
    summarize(
        nPer100People = 100 *sum(PB_nTests, na.rm = TRUE) / population_blkGrpE[1]
    ) %>% 
    left_join(block_groups(state = "IL", county = "Cook", progress_bar = FALSE), join_by(GEOID_blkGrp == GEOID)) %>%
    st_as_sf() %$%
        {leaflet(.) %>%
        addProviderTiles(providers$CartoDB.Positron) %>% 
        addPolygons(
            fillColor = ~ colorNumeric("Blues", nPer100People)(nPer100People),
            fillOpacity = 0.7,
            weight = 0,
            highlightOptions = highlightOptions(
                color = "Black",
                fillOpacity = 1,
                bringToFront = T),
            label = str_glue_data(., "<strong>{GEOID_blkGrp}</strong><br>Tests per 100 People: {round(nPer100People,2)}<br/>") %>% 
                    map(htmltools::HTML),
            labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "12px",
                direction = "auto")
        ) %>%
        addLegend(
            pal = colorNumeric("Blues", nPer100People),
            values = nPer100People,
            opacity = 0.7,
            title = "Tests per 100 People <br>By Block Group",
            position = "topright",
            na.label = "Insufficient Data"
        )} %>%
        htmlwidgets::prependContent(htmltools::tags$style(type = "text/css", "div.info.legend.leaflet-control br {clear: both;}"))
```

## Tests Per Person by Tract
```{r}
#| fig-height: 10

data %>%
    group_by(GEOID_tract) %>%
    summarize(
        nPer100People = 100 *sum(PB_nTests, na.rm = TRUE) / population_tractE[1]
    ) %>% 
    left_join(tracts(state = "IL", county = "Cook", progress_bar = FALSE), join_by(GEOID_tract == GEOID)) %>%
    st_as_sf() %$%
        {leaflet(.) %>%
        addProviderTiles(providers$CartoDB.Positron) %>% 
        addPolygons(
            fillColor = ~ colorNumeric("Blues", nPer100People)(nPer100People),
            fillOpacity = 0.7,
            weight = 0,
            highlightOptions = highlightOptions(
                color = "Black",
                fillOpacity = 1,
                bringToFront = T),
            label = str_glue_data(., "<strong>{GEOID_tract}</strong><br>Tests per 100 People: {round(nPer100People,2)}<br/>") %>% 
                    map(htmltools::HTML),
            labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "12px",
                direction = "auto")
        ) %>%
        addLegend(
            pal = colorNumeric("Blues", nPer100People),
            values = nPer100People,
            opacity = 0.7,
            title = "Tests per 100 People <br>By Tract",
            position = "topright",
            na.label = "Insufficient Data"
        )} %>%
        htmlwidgets::prependContent(htmltools::tags$style(type = "text/css", "div.info.legend.leaflet-control br {clear: both;}"))
```
:::

::: panel-tabset

## Positive Test Percentage by Mean Building Age
```{r}
data %>%
    group_by(GEOID_tract) %>%
    summarize(
        bldAge_mean = sum(bldAge_mean * bld_number, na.rm = TRUE)/sum(bld_number, na.rm = TRUE),
        PB_gt1Pct = sum(PB_gt1Pct * PB_nTests, na.rm = TRUE)/sum(PB_nTests, na.rm = TRUE)
    ) %>% 
    ggplot(aes(x = bldAge_mean, y = PB_gt1Pct)) +
        geom_point(color = "steelblue1") +
        labs(
            title = "Mean Building Age vs Percent of Tests with Greater than 1 \nPPB Lead Detected by Tract",
            x = "Mean Building Age (years)",
            y = "Tests with Greater than 1 PPB Lead Detected (%)") +
        scale_y_continuous(labels = scales::percent) +
        theme(
            plot.title = element_text(hjust = 0.5),
            legend.position="none"
        )
```

## Positive Test Percentage by Percent Built Before 1986
```{r}
data %>%
    group_by(GEOID_tract) %>%
    summarize(
        bldAge_nBfr86 = 1-((sum(bldAge_nAft86, na.rm = TRUE))/sum(bld_number, na.rm = TRUE)),
        PB_gt1Pct = sum(PB_gt1Pct * PB_nTests, na.rm = TRUE)/sum(PB_nTests, na.rm = TRUE)
    ) %>% 
    ggplot(aes(x = bldAge_nBfr86, y = PB_gt1Pct)) +
        geom_point(color = "steelblue1") +
        labs(
            title = "Percent of Buildings built after 1986 vs Percent of Tests with \nGreater than 1 PPB Lead Detected by Tract",
            x = "Buildings built after 1986 (%)",
            y = "Tests with Greater than 1 PPB Lead Detected (%)") +
        scale_x_continuous(labels = scales::percent) +
        scale_y_continuous(labels = scales::percent) +
        theme(
            plot.title = element_text(hjust = 0.5),
            legend.position="none"
        )
```
:::

```{r}
#| fig-height: 10
#| fig-width: 10

data %>%
    select(where(is.numeric)) %>%
    select(where(function(x) sum(!is.na(x)) != 0)) %>%
    select(where(function(x) var(x, na.rm = TRUE) != 0)) %>% 
    correlate(method = "spearman") %>%
    autoplot()
```

## Modeling


```{r}
tidymodels_prefer()
```

#### Creating Test/Training Data

```{r}
set.seed(345) 
model_fitting_data <- data %>%
    mutate(PB_nTests = 
        case_when(
            is.na(PB_nTests) ~ 5,
            .default = PB_nTests
        )
    ) %>%
    filter(!is.na(PB_majoritygt1)) %>%
    mutate(PB_majoritygt1 = as.factor(PB_majoritygt1)) %>%
    filter(GEOID_CA == "22") %>%
    select(-PB_gt1Pct, -GEOID_blk, -GEOID_CA)

split <- initial_split(model_fitting_data, strata = PB_majoritygt1)
train <- training(split)
test  <- testing(split)

set.seed(456) 
resamples <- vfold_cv(train, v = 5, strata = PB_majoritygt1)
```

#### Creating Our Recipes

```{r}
default_recipe <- model_fitting_data %>%    
    recipe(PB_majoritygt1 ~ .) %>%   
    step_novel(all_nominal_predictors()) %>%   
    step_dummy(all_nominal_predictors()) %>%
    step_zv(all_predictors()) %>%
    step_smote(all_outcomes())

normalized_recipe <- default_recipe %>%
    step_normalize(all_predictors())
```

#### Creating Our Models

```{r}
logistic_class_spec <- 
    logistic_reg(penalty = tune(), mixture = tune()) %>% 
    set_engine("glmnet") %>% 
    set_mode("classification")

lightGBM_class_spec <- 
    boost_tree(tree_depth = tune(), learn_rate = tune(), loss_reduction = tune(), min_n = tune(), sample_size = tune(), trees = tune()) %>%
    set_engine("lightgbm") %>%
    set_mode("classification")

rf_class_spec <- 
  rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
  set_engine("ranger") %>% 
  set_mode("classification")
```

#### Creating Workflow

```{r}
normalized <- workflow_set(
    preproc = list(normalized = normalized_recipe),
    models = list(logistic = logistic_class_spec))

no_pre_proc <- workflow_set(
    preproc = list(simple = default_recipe), 
    models = list(RF = rf_class_spec, boosting = lightGBM_class_spec))
```

```{r}
all_workflows <- bind_rows(no_pre_proc, normalized)
```

#### Tune Model

```{r}
conflicted::conflicts_prefer(purrr::flatten)
conflicted::conflicts_prefer(purrr::set_names)

grid_results <- all_workflows %>%
  workflow_map(
    "tune_race_anova",
    seed = 567,
    resamples = resamples,
    grid = 30,
    control = control_race(
      save_pred = FALSE,
      save_workflow = FALSE,
      parallel_over = "everything"))
```

```{r}
autoplot(grid_results)
```

#### Select Best Model

```{r}
conflicted::conflicts_prefer(dplyr::filter)

best_wf <- grid_results %>% 
    rank_results() %>%
    filter(.metric == "roc_auc", rank == 1) %>%
    pull(wflow_id)

best_results <- grid_results %>%
    extract_workflow_set_result(best_wf) %>% 
    select_best(metric = "roc_auc") 

best_results_fit <- 
    grid_results %>% 
    extract_workflow(best_wf) %>%
    finalize_workflow(best_results) %>% 
    last_fit(split = split)
```


```{r}
bayes_results <- grid_results %>% 
    extract_workflow(best_wf) %>%
    tune_bayes(
        resamples = resamples,
        initial = 20,
        iter = 25,
        control = control_bayes(verbose = TRUE)
    )
```


```{r}
bayes_results %>%
  show_best(metric = "roc_auc") %>% 
  knitr::kable()
```


```{r}
final_fit <- grid_results %>% 
    extract_workflow(best_wf) %>%
    finalize_workflow(select_best(bayes_results, metric = "roc_auc")) %>% 
    last_fit(split = split)
```


```{r}
final_fit %>%
    collect_predictions() %>%
    roc_curve(PB_majoritygt1, .pred_FALSE) %>%
    autoplot()
```

```{r}
final_fit %>% 
    collect_metrics() %>%
    knitr::kable()
```


```{r}
final_fit %>%
    collect_predictions() %>%
    conf_mat(PB_majoritygt1, .pred_class) %>%
    autoplot(type = "heatmap")
```